\LaTeXe\ guide for the Journal of Functional Programming

### 2019-05-02 papers/jfp/7-experience.tex
Improve things that were confusing to SPJ in S7

for various choices of~\hs{c}, are studied in depth by~Jaskelioff and
O’Connor~\shortcite{jaskelioff2015representation}.

@@

Converting to a build suitable for sharing is
not overly onerous, but nor is it trivial. Some things that require attention:

\begin{itemize}
\item \textbf{Irrelevant differences}: A common problem is that you do not get
      shared caching when you want it. As one example, imagine two users install
      \cmd{gcc} on different paths (say \cmd{/usr/bin/gcc} and \cmd{/usr/local/bin/gcc}).
      If these paths are recorded by the build system, the users won't share
      cache entries. As another example, consider a compiler that embeds the current time
      in the output: any users who build that file locally won't get any shared
      caching of subsequent outputs. Possible solutions include using relative paths;
      depending only on version numbers for system binaries (e.g. \cmd{gcc});
      controlling the environment closely (e.g. using \Nix); and extra flags to
      encourage compilers to be more deterministic.
\item \textbf{Insufficient produced files}: A build rule must declare all files
      it produces, so these can be included in the cache.
      As an example using Haskell, compilation of \cmd{Foo.hs} produces
      \cmd{Foo.hi} and \cmd{Foo.o}. If you declare the rule as producing \cmd{Foo.hi},
      and other rules \emph{depend on} \cmd{Foo.hi}, but \emph{also use} \cmd{Foo.o} after
      depending on \cmd{Foo.hi}, a local build will probably work. However,
      if \cmd{Foo.hi} is downloaded from a remote cache, \cmd{Foo.o} will not be present,
      and subsequent commands may fail (e.g. linking). In practice, most issues encountered
      during the move to cloud builds for GHC were caused by failing to declare produced
      files.
\item \textbf{Missing dependencies}: While missing dependencies are
      always a problem, the move to a cloud build makes them even more serious.
      Suppose you omit a dependency on a file \cmd{foo} that is produced by a rule,
      but that changes rarely (e.g. perhaps \cmd{foo} is derived from
      a configuration input file).
      With purely local builds, each developer will build \cmd{foo} at least once (when
      starting from scratch), and more regularly if they ever do clean
      builds. In contrast, with a cloud build, the file might only be built
      \emph{just once}, and reused forevermore. If things go wrong, a clean build
      is no longer a refuge of last resort.
\end{itemize}

@@

A small number of GHC build rules
cannot be cached since they mutate the bundled package database (e.g. to
register a new library); \Cloud~\Shake provides a way to manually label such
build rules, thereby excluding them from caching.

>>>>

\noindent for various choices of~\hs{c}, are studied in depth by~Jaskelioff and
O’Connor~\shortcite{jaskelioff2015representation}.

@@

Converting to a build suitable for sharing is not overly onerous, but nor is it trivial.
In particular, a cloud build is less forgiving about untracked operations -- things that
are wrong but usually harmless in local builds often cause serious problems in a cloud setting.
Some things that require attention in moving to a cloud build:

\begin{itemize}
\item \textbf{Irrelevant differences}: A common problem is that you do not get
      shared caching when you want it. As one example, imagine two users install
      \cmd{gcc} on different paths (say \cmd{/usr/bin/gcc} and \cmd{/usr/local/bin/gcc}).
      If these paths are recorded by the build system, the users won't share
      cache entries. As another example, consider a compiler that embeds the current time
      in the output: any users who build that file locally won't get any shared
      caching of subsequent outputs. Possible solutions include using relative paths;
      depending only on version numbers for system binaries (e.g. \cmd{gcc});
      controlling the environment closely (e.g. using \Nix); and extra flags to
      encourage compilers to be more deterministic.
\item \textbf{Insufficient produced files}: A build rule must declare all files
      it produces, so these can be included in the cache.
      As an example using Haskell, compilation of \cmd{Foo.hs} produces
      \cmd{Foo.hi} and \cmd{Foo.o}. If you declare the rule as producing \cmd{Foo.hi},
      and other rules \emph{depend on} \cmd{Foo.hi}, but \emph{also use} \cmd{Foo.o} after
      depending on \cmd{Foo.hi}, a local build will probably work (although treating \cmd{Foo.o}
      as a proper dependency would definitely be preferable). However,
      if \cmd{Foo.hi} is downloaded from a remote cache, \cmd{Foo.o} will not be present,
      and subsequent commands may fail (e.g. linking).
      In practice, most issues encountered
      during the move to cloud builds for GHC were caused by failing to declare produced
      files.
\item \textbf{Missing dependencies}: While missing dependencies are
      always a problem, the move to a cloud build makes them more serious.
      With local builds outputs will be built at least once per user, but with a
      cloud build they might only be built \emph{once ever}.
\end{itemize}

@@

A small number of GHC build rules cannot be cached. These rules register new libraries
in the GHC package database, and rather than producing a output file, they mutate
a shared file. \Cloud~\Shake provides a way to manually label such
build rules, thereby excluding them from caching.
